---
title: "Chi-Squared"
format: 
  html: 
    self-contained: true
    toc: true             
  pdf: 
    number-sections: true 
    toc: true            
    pdf-engine: xelatex   
editor: visual
---

```{r}
library(ggplot2)
library(descr)
```

## Example using the Iris Data

Alternative datasets:

-   mtcars:

    -   cyl (Cylinders: 4, 6, 8)

    -   am (Transmission: 0 = automatic, 1 = manual)

    -   gear (Number of gears)

-   warpbreaks

    -   wool (A or B)

    -   tension (L, M, H)

### Load Data

```{r}
data(iris)
str(iris)
```

### Data Manipulation

```{r}
# View(iris)
```

#### Species

```{r}
freq(as.ordered(iris$Species), plot = FALSE)
```

#### Sepal Length

```{r}
freq(as.ordered(iris$Sepal.Length), plot = FALSE)
```

```{r}
# Create a categorical version of Sepal.Length
iris$SepalCat <- cut(iris$Sepal.Length,
                     breaks = 3,
                     labels = c("Short", "Medium", "Long"))
```

```{r}
freq(as.ordered(iris$SepalCat), plot = FALSE)
```

### Create Table

```{r}
# Table with Speal Category and Species
table_iris <- table(iris$SepalCat, iris$Species)
print(table_iris)
```

### Assumptions

```{r}
# Cells should be greater than 5
chisq.test(table_iris)$expected
```

### Chi-squared Test

```{r}
# Cells have less than 5, Fisher's test would be appropriate
# Will proceed with Pearson's to capture test statistic
chisq_test_result <- chisq.test(table_iris)
chisq_test_result
```

**Interpretation:** A chi-square test of independence was conducted to determine if there was an association between species (setosa, versicolor, virginica) and sepal length (short, medium, or long). The results were significant, $\chi^2$ (`r chisq_test_result$parameter`)= `r round(chisq_test_result$statistic, 2)`, *p* \< `r max(round(chisq_test_result$p.value, 3), 0.001)`.
