---
title: "dataManagement"
format: 
  html: 
    self-contained: true
editor: visual
---

# Introduction

Dataset: \[Name of the Data that you are using\]

Source: \[Briefy describe the source of your data\]

Variables of Interest:

-   'Var1': \[Variable description\]

-   'Var2': \[Variable description\]

## Load libraries

```{r}
library(dplyr)
```

## Load data

```{r}
# Define the file path (going up one level from /scripts to the main directory)
file_path <- file.path("..", "data", "data.csv")

# Load the CSV file into a dataframe (automatically a tibble)
df <- read_csv(file_path)

# View the first few rows
head(df)
```

## Subsetting the Data

Limit the variables

```{r}
# List of variables to keep
selected_vars <- c("unique_id", "survey_id", "country", "language", "incentive", 
                   "age", "gender", "degree_major", "degree_level_maths",
                   "Q7.1_1", "Q7.1_2", "Q7.1_3", "Q7.1_4", "Q7.1_5", "Q7.1_6", "Q7.1_7", "Q7.1_8", "Q7.1_9", "Q7.1_10",
                   "Q8.1_1", "Q8.1_2", "Q8.1_3", "Q8.1_4", "Q8.1_5", "Q8.1_6", "Q8.1_7", "Q8.1_8", "Q8.1_9", "Q8.1_10",
                   "Q9.1_1", "Q9.1_2", "Q9.1_3", "Q9.1_4", "Q9.1_5", "Q9.1_6", "Q9.1_7", "Q9.1_8", "Q9.1_9", "Q9.1_10",
                   "Q10.1_1", "Q10.1_2", "Q10.1_3", "Q10.1_4", "Q10.1_5",
                   "Q11.1_1", "Q11.1_2", "Q11.1_3", "Q11.1_4",
                   "Q12.1_1", "Q12.1_2", "Q12.1_3", "Q12.1_4",
                   "Q13.1_1", "Q13.1_2", "Q13.1_3", "Q13.1_4",
                   "Q14.1_1", "Q14.1_2", "Q14.1_3", "Q14.1_4",
                   "Q15.1_1", "Q15.1_2", "Q15.1_3", "Q15.1_4",
                   "Q16.1_1", "Q16.1_2", "Q16.1_3", "Q16.1_4",
                   "Q17.1", "Q17.2", "Q17.3")


# Create focused data frame with only variables that I might be using.
dfSubset <- df |>
  select(all_of(selected_vars))

head(dfSubset)
```

# Identifying and Handling Missing Data

Invalid values must be replaced with NA. The summary() function can help find variables that need to be modified.

```{r}
summary(df)
```
