---
title: "Univariate Data Visualization"
format: 
  html: 
    self-contained: true
    toc: true             
  pdf: 
    number-sections: true 
    toc: true            
    pdf-engine: xelatex 
editor: visual
---

## Univariate Data Visualizations

```{r}
library(ggplot2)
library(dplyr)
library(papaja)
```

```{r}
#| echo: false
#| message: false
#| warning: false
source("loadFilteredData.R")
```

```{r}
# View(df)
```

### Figure 1 Bar Graph

```{r}
# univariate graph (categorical variable)
df %>% 
  ggplot(aes(x=age))+
  geom_bar()+
  ggtitle("Age of Participants")
```

```{r}
# univariate graph (categorical variable)
df %>% 
  ggplot(aes(x=age)) +
  geom_bar() +
  ggtitle("Age of Participants") +
  xlab("Age") +
  ylab("Frequency") +
  theme_apa() # from papaja package
```

The majority of the participants in the sample are between the ages of 18-21, which is typical for students in college. The sample includes non-traditional students as well, but not as many.

### Figure 2 Histogram

```{r}
# univariate graph (categorical variable)
df %>% 
  ggplot(aes(x=BFNE_Total)) +
  geom_histogram() +
  ggtitle("Fear of Negative Evaluation")
```

```{r}
# univariate graph (categorical variable)
df %>% 
  filter(!is.na(BFNE_Total)) %>% 
  ggplot(aes(x=BFNE_Total)) +
  geom_histogram(binwidth = 5, fill = "lightblue", color = "black") +
  ggtitle("Fear of Negative Evaluation") +
  theme_apa()
```

This variable represents fear of negative evaluation. Scores for this variable range from 1 to 40. The data is slightly skewed to the left. This suggests that participants had slightly higher reporting of fear of negative evaluation.
