# Folder setup

Save raw data file in new folder called /data/raw

Save another copy of the data as "data" in /data folder.

# File setup

Install the required packages.

```{r}
install.packages("tidyverse")
```

```{r}
install.packages("descr")
```

Load the required packages.

```{r}
# Load necessary libraries
library(tidyverse)  # For data manipulation and tibbles
library(descr) # For frequency tables
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
```

Hide messages with #\| echo: false

Load data file into path

```{r}
# Define the file path (going up one level from /scripts to the main directory)
file_path <- file.path("..", "data", "data.csv")

# Load the CSV file into a dataframe (automatically a tibble)
df <- read_csv(file_path)

# View the first few rows
head(df)
```

# View data

Check the data after it has been loaded.

Check the column names.

```{r}
# Check column names
colnames(df)

```

View summary statistics

```{r}
# View summary statistics
summary(df)
```

Check the variable types.

```{r}
# Check data structure
glimpse(df)
```

# Create Frequency Tables

Use descr package to create frequency tables.

```{r}
# Frequency table for country where data was collected
freq(as.ordered(df$country), plot = FALSE)
```

**Interpretation:** While I am from the US, the sample size from the US is small. I will likely use country as an additional variable or limit the analyses to certain countries (TBD based on additional exploration).

```{r}
# Frequency table for language spoken for participants
freq(as.ordered(df$language), plot = FALSE)
```

**Interpretation:** There is a large sample of English speakers. I will focus my analyses on English speakers since English is my primary language. From my readings, I know that language can influence individuals processing of math. This limitation to English speakers will help focus my analysis and eliminate variation due to language.

```{r}
# Frequency table for country where data was collected
freq(as.ordered(df$degree_major), plot = FALSE)
```

**Interpretation:** Most of participants were psychology majors. I would like to focus my analyses on psychology majors since this is a Statistics class for psychology majors.

# Extra Additional Table

Choosing one additional variable to format differently.

```{r}
install.packages("knitr")
install.packages("dplyr")
```

```{r}
library(knitr)
library(dplyr)
```

```{r}
# Create the data for the table
genderTable <- as.data.frame(freq(as.ordered(df$gender), plot = FALSE))

genderTable
```

```{r}
# Create columns for new table
colnames(genderTable)
```
